---
title: "How to use UNHCR themes?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use UNHCR themes?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

## Set Up: load packages

```{r message=FALSE, warning=FALSE}
library(unhcRstyle)
library(ggplot2)
library(tidyverse)
library(unhcrdatapackage)
update_geom_font_defaults()


```


## Use Theme on a bar plot

To understand this theme, let's first consider a basic ggplot.

```{r}
# make fake data for plots ----
bar <- count(mpg, class)

## chart
g_bars <- ggplot(bar, aes(class, n)) + 
          geom_col( )  +
          geom_text(aes(label=n), nudge_y = 3) +
            labs(title="Title: Main Message supported by the chart (less than 80 char)",
                 subtitle="Subtitle: Interpretation on what is displayed",
                 x="Title for X axis - limit to less than 40 char", 
                 y="Title for X axis",
                 caption="Source of data / Reference / Disclaimer")
g_bars
```


We can now:
 * reorder x based on frequency to make the chart more legible
 * flip the chart again to make the chart more legiblel
 * include the standard `unhcr_blue`
 * apply the UNHCR them with `unhcr_theme()` 

```{r}
g_bars <- ggplot(bar, aes(x= reorder(class, n), y = n)) +
          geom_col( fill = unhcr_blue)  +
          geom_text(aes(label=n), nudge_y = 3) +
          coord_flip() +
            labs(title="Title: Main Message supported by the chart (less than 80 char)",
                 subtitle="Subtitle: Interpretation on what is displayed",
                 x="Title for X axis - limit to less than 40 char", 
                 y="Title for X axis",
                 caption="Source of data / Reference / Disclaimer") +
          unhcr_theme(grid="Y")

g_bars
```

The main title are directly aligned, the correct font is used.


## Use Theme on a scatter plot

To understand this theme, let's first consider a basic ggplot.

```{r}
# make fake data for plots ----
points <- data.frame(x = 1:10, y = 1:10)

## chart
g_points <- ggplot(points, aes(x, y, col = factor(x))) + 
            geom_point(size = 5) +
            labs(x="Title for X axis", 
                 y="Title for X axis",
                 title="Main Message supported by the chart (less than 80 char)",
                 subtitle="Interpretation on what is displayed",
                 caption="Source of data / Reference")
g_points
```


We can now apply the UNHCR them with `unhcr_theme()`

```{r}
g_points <- g_points + 
            unhcr_theme()

g_points
```

The main title are directly aligned, the correct font is used.

next step is to apply one UNHCR color palette.

```{r}
g_points <- g_points + 
  scale_color_discrete_unhcr(palette = "test", extend = TRUE) 
  #scale_color_discrete_unhcr(palette = "unhcr", extend = TRUE) 


g_points
```

