---
title: "How to use UNHCR themes?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use UNHCR themes?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>",
  fig.path = "figures/charts-",
  out.width = "100%"
)
```

The [unhcRstyle](https://github.com/UNHCR-WEB/unhcRstyle) package sets the basic the visual elements that compose the UNHCR Brand. It includes some color palettes as well as a ggplot2 theme to help you create UNHCR branded charts and reports.

## Set Up: load packages

```{r message=FALSE, warning=FALSE}
library(unhcRstyle)
library(ggplot2)
library(tidyverse)
library(unhcrdatapackage)
library(scales)
library(ragg)
update_geom_font_defaults()
```

## Getting the data

Use of the [unhcrdatapackage](https://github.com/Edouard-Legoupil/unhcrdatapackage/) to get End-year population figures - these are stock figures for specific types of populations at the end of each year, including refugees, IDPs and asylum seekers.

```{r message=FALSE, warning=FALSE}
popdata <- unhcrdatapackage::end_year_population_totals
```

## Line charts

Use of [unhcr_theme]() and data from [unhcrdatapackage](https://github.com/Edouard-Legoupil/unhcrdatapackage/) to create UNHCR branded line charts.

### Single line chart

Simplify the data to have only the total refugee population at the end of each year.

```{r message=FALSE, warning=FALSE}
ref_year <- popdata %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarise(REF = sum(REF, na.rm = TRUE)) %>% 
  dplyr::mutate(Year = as.Date(Year, format = '%Y'))
```

Plot the simplify data using in a single line chart using [unhcr_theme]().

```{r single-line, message=FALSE, warning=FALSE}
# Create plot
single_line <- ggplot2::ggplot(ref_year, aes(x = Year, y = REF)) + #Assign Year to x and REF population value to y
  geom_line(colour = unhcr_blue, size = 1) + #Define line chart color to unhcr_blue
  geom_hline(yintercept = 0, size = 1, colour = unhcr_grey) + #Add major line at 0
  labs(title = "Total number of refugees",
       subtitle = "Evolution of refugee population 1951-2020",
       caption = "Source: UNHCR Refugee Statistics") +
  scale_y_continuous(label = scales::label_number_si()) + ## Format axis number
  scale_x_date(breaks = breaks_pretty(n = 10)) + #Add axis label every 10 years
  unhcr_theme(base_size = 16, grid = "Y") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) #Remove axis title in this context

# Save plot
pngfile <- knitr::fig_path(".png") # Set the file to use the Rmd chunk name as file name
agg_png(pngfile, width = 1280, height = 960, units = "px", res = 144) # Define plot size
plot(single_line)
invisible(dev.off())
knitr::include_graphics(pngfile) #Include graphic while knitting
```

### Multiple lines

Simplify the data to have only the total refugee and idp population at the end of each year.

```{r message=FALSE, warning=FALSE}
ref_idp_year <- popdata %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarise(REF = sum(REF, na.rm = TRUE),
                   IDP = sum(IDP, na.rm = TRUE)) %>%
  tidyr::pivot_longer(cols = c(REF, IDP), names_to = "population_type", values_to = "number_people") %>% 
  dplyr::mutate(Year = as.Date(Year, format = '%Y'))
```

Plot the simplify data using in a multiple lines chart using [unhcr_theme]().

```{r multi-line, message=FALSE, warning=FALSE}
# Create plot
multi_line <- ggplot2::ggplot(ref_idp_year, aes(x = Year, y = number_people, colour = population_type)) + #Assign Year to x and population value to y
  geom_line(size = 1) + #Define line chart color to the population type
  geom_hline(yintercept = 0, size = 1, colour = unhcr_grey) + #Add major line at 0
  labs(title = "Total number of refugees and IDPs",
       subtitle = "Evolution of refugee and IDP population from 1951 to 2020",
       caption = "Source: UNHCR Refugee Statistics") +
  ylab("Persons") +
  scale_y_continuous(label = scales::label_number_si()) + ## Format axis number
  scale_x_date(breaks = breaks_pretty(n = 10)) + #Add axis label every 10 years
  scale_color_manual(labels = c("IDPs", "Refugees"),
                     values = usecol(pal = c(unhcr_green, unhcr_blue))) + #Set color from the unhcr_pal
  unhcr_theme(base_size = 16, grid = "Y") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

# Save plot
pngfile <- knitr::fig_path(".png") # Set the file to use the Rmd chunk name as file name
agg_png(pngfile, width = 1280, height = 960, units = "px", res = 144) # Define plot size
plot(multi_line)
invisible(dev.off())
knitr::include_graphics(pngfile) #Include graphic while knitting
```

## Use Theme on a bar plot

To understand this theme, let's first consider a basic ggplot.

```{r}
# make fake data for plots ----
bar <- count(mpg, class)

## chart
g_bars <- ggplot(bar, aes(class, n)) + 
          geom_col( )  +
          geom_text(aes(label=n), nudge_y = 3) +
            labs(title="Title: Main Message supported by the chart (less than 80 char)",
                 subtitle="Subtitle: Interpretation on what is displayed",
                 x="Title for X axis - limit to less than 40 char", 
                 y="Title for X axis",
                 caption="Source of data / Reference / Disclaimer")
g_bars
```


We can now:
 * reorder x based on frequency to make the chart more legible
 * flip the chart again to make the chart more legiblel
 * include the standard `unhcr_blue`
 * apply the UNHCR them with `unhcr_theme()` 

```{r}
g_bars <- ggplot(bar, aes(x= reorder(class, n), y = n)) +
          geom_col( fill = unhcr_blue)  +
          geom_text(aes(label=n), nudge_y = 3) +
          coord_flip() +
            labs(title="Title: Main Message supported by the chart (less than 80 char)",
                 subtitle="Subtitle: Interpretation on what is displayed",
                 x="Title for X axis - limit to less than 40 char", 
                 y="Title for X axis",
                 caption="Source of data / Reference / Disclaimer") +
          unhcr_theme(grid="Y")

g_bars
```

The main title are directly aligned, the correct font is used.


## Use Theme on a scatter plot

To understand this theme, let's first consider a basic ggplot.

```{r}
# make fake data for plots ----
points <- data.frame(x = 1:10, y = 1:10)

## chart
g_points <- ggplot(points, aes(x, y, col = factor(x))) + 
            geom_point(size = 5) +
            labs(x="Title for X axis", 
                 y="Title for X axis",
                 title="Main Message supported by the chart (less than 80 char)",
                 subtitle="Interpretation on what is displayed",
                 caption="Source of data / Reference")
g_points
```


We can now apply the UNHCR them with `unhcr_theme()`

```{r}
g_points <- g_points + 
            unhcr_theme()

g_points
```

The main title are directly aligned, the correct font is used.

next step is to apply one UNHCR color palette.

```{r}
g_points <- g_points + 
  scale_color_discrete_unhcr(palette = "test", extend = TRUE) 
  #scale_color_discrete_unhcr(palette = "unhcr", extend = TRUE) 


g_points
```

